---
title: "Officer Package Demo"
author: "Libby Heeren"
date: "2/26/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(formattable)
library(officer)
```

This  demo would'nt be possible without the guidance and mentorship of my former director, Walt DeGrange. He taught me all I know about the officeverse. Thanks, Walt!  

The below code chunk is a formatting chunk, the yInfl sets the y axis to be 1.25 times larger than the max y value, allowing headroom for data callouts on chart bars. I've set the colors using the tropical palette from my coolers.co collection: https://coolors.co/palette/e08e94-e5ce6e-dadc89-bfe8d6-5ac1be-66a5c4-7973b3-d6d6d6  

For the colorBlindFriendly setting, if set to TRUE, ggplots will use the viridis palette, which doesn't rely on red/green contrast.  

```{r formatting, include = FALSE}

# Set font size for axis text (tick marks)
atx = 11
aty = 11


# Set font size for axis labels
alx = 15
aly = 15


#Set y-axis inflation factor
yInfl <- 1.25


```

This is the chunk where the template Word and PowerPoint files are read in. For my uses in this demo, I will use the standard templates that the officer package author David Gohel uses in his github (https://github.com/davidgohel/officer), but with some modifications of my own to the styles by choosing a different theme and adding a title and TOC (table of contents) page.   


```{r Import Report and Presentation}


# Report will be the Word document, Presentation will be the PowerPoint deck

# These objects will contain the blank template file which we will add to by creating objects called ReportNew and PresentationNew. 


Report <- read_docx("template.docx")


Presentation <- read_pptx("template.pptx")
```

Let's import some data. The Facebook data set has 490 rows and 19 columns. We will remove all rows with missing data. It can be found here: https://www.kaggle.com/roxannac/facebook-data  

```{r Import and Modify Data}

# read in facebook data set
facebook = read.csv("dataset_Facebook.csv", header=TRUE)

# remove all rows with missing values
fbook = facebook[complete.cases(facebook),]

# remove all output variables except Total.Interactions 
# and Lifetime.Post.Total.Reach
removedVars <- c(9:18)
fbook2 <- fbook[, -removedVars]

# define "month" (numeric month is column 4 in the facebook data set)
fbook2$Month <- month.abb[fbook2$Post.Month]

# reorder month to be an ordered factor so months aren't in alphabetical order
fbook2$Month <-factor(fbook2$Month, 
                      levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# define "weekday" (numeric weekday is column 5 in the facebook data set)
weekday=c()
weekday[fbook2[,5]==2]="Mon"; weekday[fbook2[,5]==3]="Tue"
weekday[fbook2[,5]==4]="Wed"; weekday[fbook2[,5]==5]="Thu"
weekday[fbook2[,5]==6]="Fri"; weekday[fbook2[,5]==7]="Sat"; 
weekday[fbook2[,5]==1]="Sun"

# add weekday variable to dataset
fbook2$Weekday <- weekday

# reorder weekday to be an ordered factor so days aren't in alphabetical order
fbook2$Weekday <-factor(fbook2$Weekday, 
                      levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat",
                                 "Sun"))

# create a factor variable for paid vs unpaid posts (column 7)
paid=c()
paid[fbook2[,7]==1]="Paid"; paid[fbook2[,7]==0]="Unpaid"

fbook2$Paid_or_Unpaid <- paid
```

Now that we have our data in place and ready to report on, we can begin adding pages to our document. Let's start by exploring Total.Interaction by month.  

```{r average Total.Interactions by month, echo=FALSE, message=FALSE, warning=FALSE}

# Create a data table containing average total interactions by month

TotalInteraction <- fbook2 %>% 
  select(Total.Interactions, Month) %>%
  group_by(Month) %>%
  summarise(Average = comma(mean(Total.Interactions),0)) %>% # formattable::comma
  ungroup()

print(TotalInteraction)

# Create bar charts to use in word (tall) and powerpoint (wide)

# Set y axis limit to be 1.25 times larger than the max average value
ylim = max(TotalInteraction$Average) * yInfl # yInfl = 1.25, set above

# Create a tall chart 

TotalInteractionChart_tall <- 
  ggplot(data = TotalInteraction, mapping = aes(x = Month, y = Average)) + 
  stat_summary(fun.data = mean_sdl, geom = "bar") +
  scale_x_discrete() +
  geom_text(aes(label = Average),
  hjust = 0.5, vjust = -0.3 , nudge_x = 0.0) +
  # xlab("") +
  ylab("Average Total Post Interactions") +
  theme_minimal()+
  theme(axis.text.x = element_text(size=atx),
        axis.text.y = element_text(size=aty),
        axis.title.x = element_text(size=alx),
        axis.title.y = element_text(size=aly)) +
  scale_y_continuous(limits = c(0, ylim))

print(TotalInteractionChart_tall)
  
# Create a wide chart by using coord_flip()

TotalInteractionChart_wide <- 
  ggplot(data = TotalInteraction, mapping = aes(x = Month, y = Average)) + 
  stat_summary(fun.data = mean_sdl, geom = "bar") +
  scale_x_discrete(limits = rev(levels(TotalInteraction$Month))) +
  geom_text(aes(label = Average),
  hjust = 0.0, vjust = 0.5 , nudge_x = 0.0) +
  coord_flip() +
  # xlab("") +
  ylab("Average Total Post Interactions") +
  theme_minimal()+
  theme(axis.text.x = element_text(size=atx),
        axis.text.y = element_text(size=aty),
        axis.title.x = element_text(size=alx),
        axis.title.y = element_text(size=aly)) +
  scale_y_continuous(limits = c(0, ylim))

print(TotalInteractionChart_wide)

# Edit Parameters - if plots are the same for both docs, just use one for both ggs

H1 <- "Average Total Interactions by Month" # Page/Slide Title goes here
ggWide <- TotalInteractionChart_wide # wider ggplot object goes here
ggTall <- TotalInteractionChart_tall # taller ggplot object goes here
TableTitle <- "Average Total Interactions Table" # Data Table title goes here
DT <- TotalInteraction # Data table object goes here


# Long form text parameters

dataDescription <- 
  "This section shows the average Total Interactions on posts by month and includes both paid (paid facebook to promote the post) and unpaid posts. In the plot above, we see that February, May, and September show the highest average levels of Total.Interactions per post."


dataStrengths <- 
  "Data strengths go here, delete sections as needed."


dataLimitations <- 
  "Data limitations go here."


dataAssumptions <- 
  "Data assumptions go here."




# Add a page to the Word report doc: Title, chart, table, desc., limits, assmps

# DO NOT EDIT THIS, the parameters above will be used

# Notice we are not modifying Report or Presentation, we are making new ones

ReportNew <- Report %>%
  body_add_break() %>% # Adds a page break so this starts on a new page
  body_add_par(value = H1, style = "heading 1") %>% # Title
  body_add_gg(value = ggTall, style = "centered") %>% # ggplot
  body_add_par(value = "") %>% # Adds a space between plot and table
  body_add_par(value = TableTitle, style = "heading 2") %>% # Table title
  body_add_table(value = DT, style = "table_template") %>% # Data table
  body_add_par(value = "") %>% # Adds a space between table and text
  body_add_par(value = "Description", style = "heading 2") %>% 
  body_add_par(value = dataDescription, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Strengths", style = "heading 2") %>% 
  body_add_par(value = dataStrengths, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Limitations", style = "heading 2") %>% 
  body_add_par(value = dataLimitations, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Assumptions", style = "heading 2") %>% 
  body_add_par(value = dataAssumptions, style = "Normal")
  
print(ReportNew, target = "Demo_Report.docx")


# Add another slide to the pptx file: Title and graph/chart
# DO NOT EDIT THIS, the parameters above will be used
PresentationNew <- Presentation %>%
  add_slide(layout = "Title and Content", master = "Facet") %>% 
  ph_with(value = H1, location = ph_location_type(type = "title")) %>% 
  ph_with(ggWide, location = ph_location_type(type = "body"))


print(PresentationNew, target = "Demo_Deck.pptx")


```

Now, let's add a section exploring the average Total Interactions on posts by Weekday. All we have to do is copy the chunk above, change a few parameters, change the text descriptions as needed, and we're done. We will change the Month variable to Weekday, and change any titles containing "by Month" to "by Week," as well as the description of our findings.    

```{r average Total.Interactions by weekday, echo=FALSE, message=FALSE, warning=FALSE}

# Create a data table containing average total interactions by week

TotalInteraction <- fbook2 %>% 
  select(Total.Interactions, Weekday) %>%
  group_by(Weekday) %>%
  summarise(Average = comma(mean(Total.Interactions),0)) %>% # formattable::comma
  ungroup()

# Create bar charts to use in word (tall) and powerpoint (wide)

# Set y axis limit to be 1.25 times larger than the max average value
ylim = max(TotalInteraction$Average) * yInfl # yInfl = 1.25, set above

# Create a tall chart 

TotalInteractionChart_tall <- 
  ggplot(data = TotalInteraction, mapping = aes(x = Weekday, y = Average)) + 
  stat_summary(fun.data = mean_sdl, geom = "bar") +
  scale_x_discrete() +
  geom_text(aes(label = Average),
  hjust = 0.5, vjust = -0.3 , nudge_x = 0.0) +
  # xlab("") +
  ylab("Average Total Post Interactions") +
  theme_minimal()+
  theme(axis.text.x = element_text(size=atx),
        axis.text.y = element_text(size=aty),
        axis.title.x = element_text(size=alx),
        axis.title.y = element_text(size=aly)) +
  scale_y_continuous(limits = c(0, ylim))
  
# Create a wide chart by using coord_flip()

TotalInteractionChart_wide <- 
  ggplot(data = TotalInteraction, mapping = aes(x = Weekday, y = Average)) + 
  stat_summary(fun.data = mean_sdl, geom = "bar") +
  scale_x_discrete(limits = rev(levels(TotalInteraction$Weekday))) +
  geom_text(aes(label = Average),
  hjust = 0.0, vjust = 0.5 , nudge_x = 0.0) +
  coord_flip() +
  # xlab("") +
  ylab("Average Total Post Interactions") +
  theme_minimal()+
  theme(axis.text.x = element_text(size=atx),
        axis.text.y = element_text(size=aty),
        axis.title.x = element_text(size=alx),
        axis.title.y = element_text(size=aly)) +
  scale_y_continuous(limits = c(0, ylim))



# Edit Parameters - if plots are the same for both docs, just use one for both ggs

H1 <- "Average Total Interactions by Weekday" # Page/Slide Title goes here
ggWide <- TotalInteractionChart_wide # wider ggplot object goes here
ggTall <- TotalInteractionChart_tall # taller ggplot object goes here
TableTitle <- "Average Total Interactions Table" # Data Table title goes here
DT <- TotalInteraction # Data table object goes here


# Long form text parameters

dataDescription <- 
  "This section shows the average Total Interactions on posts by Weekday and includes both paid (paid facebook to promote the post) and unpaid posts. In the plot above, we can see that Wednesdays and Sundays have the highest levels of interactions per post."


dataStrengths <- 
  "Data strengths go here, delete sections as needed."


dataLimitations <- 
  "Data limitations go here."


dataAssumptions <- 
  "Data assumptions go here."




# Add a page to the Word report doc: Title, chart, table, desc., limits, assmps

# DO NOT EDIT THIS, the parameters above will be used

# Notice we are not modifying Report or Presentation, we are making new ones

ReportNew <- Report %>%
  body_add_break() %>% # Adds a page break so this starts on a new page
  body_add_par(value = H1, style = "heading 1") %>% # Title
  body_add_gg(value = ggTall, style = "centered") %>% # ggplot
  body_add_par(value = "") %>% # Adds a space between plot and table
  body_add_par(value = TableTitle, style = "heading 2") %>% # Table title
  body_add_table(value = DT, style = "table_template") %>% # Data table
  body_add_par(value = "") %>% # Adds a space between table and text
  body_add_par(value = "Description", style = "heading 2") %>% 
  body_add_par(value = dataDescription, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Strengths", style = "heading 2") %>% 
  body_add_par(value = dataStrengths, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Limitations", style = "heading 2") %>% 
  body_add_par(value = dataLimitations, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Assumptions", style = "heading 2") %>% 
  body_add_par(value = dataAssumptions, style = "Normal")
  
print(ReportNew, target = "Demo_Report.docx")


# Add another slide to the pptx file: Title and graph/chart
# DO NOT EDIT THIS, the parameters above will be used
PresentationNew <- Presentation %>%
  add_slide(layout = "Title and Content", master = "Facet") %>% 
  ph_with(value = H1, location = ph_location_type(type = "title")) %>% 
  ph_with(ggWide, location = ph_location_type(type = "body"))


print(PresentationNew, target = "Demo_Deck.pptx")


```

Just for fun, let's see the difference in average Total Interactions for paid and unpaid facebook posts, just so that we have another section to view in our demo. We will change anything that used the Weekday variable to use the Paid_or_Unpaid variable.   

```{r average Total.Interactions by paid/unpaid, echo=FALSE, message=FALSE, warning=FALSE}

# Create a data table containing average total interactions by week

TotalInteraction <- fbook2 %>% 
  select(Total.Interactions, Paid_or_Unpaid) %>%
  group_by(Paid_or_Unpaid) %>%
  summarise(Average = comma(mean(Total.Interactions),0)) %>% # formattable::comma
  ungroup()

# Create bar charts to use in word (tall) and powerpoint (wide)

# Set y axis limit to be 1.25 times larger than the max average value
ylim = max(TotalInteraction$Average) * yInfl # yInfl = 1.25, set above

# Create a tall chart 

TotalInteractionChart_tall <- 
  ggplot(data = fbook2, mapping = aes(x = Paid_or_Unpaid, y = Total.Interactions, group = Paid_or_Unpaid, fill = Paid_or_Unpaid)) + 
  geom_boxplot(outlier.color = "red", show.legend = FALSE) +
  geom_text(data = aggregate(Total.Interactions ~ Paid_or_Unpaid, fbook2,
                             function(i) round(mean(i))), 
            aes(label = Total.Interactions),
            hjust = 0.5, vjust = 1.5 , nudge_x = 0.5) +
  xlab("Paid vs Unpaid Posts") +
  ylab("Total Post Interactions") +
  theme_minimal()+
  theme(axis.text.x = element_text(size=atx),
        axis.text.y = element_text(size=aty),
        axis.title.x = element_text(size=alx),
        axis.title.y = element_text(size=aly))

  
# Create a wide chart by using coord_flip()

TotalInteractionChart_wide <- 
  ggplot(data = fbook2, mapping = aes(x = Paid_or_Unpaid, y = Total.Interactions, group = Paid_or_Unpaid, fill = Paid_or_Unpaid)) + 
  geom_boxplot(outlier.color = "red", show.legend = FALSE) +
  geom_text(data = aggregate(Total.Interactions ~ Paid_or_Unpaid, fbook2,
                             function(i) round(mean(i))), 
            aes(label = Total.Interactions),
            hjust = 1.0, vjust = 1.5 , nudge_x = 0.5) +
  coord_flip() +
  xlab("Paid vs Unpaid Posts") +
  ylab("Total Post Interactions") +
  theme_minimal()+
  theme(axis.text.x = element_text(size=atx),
        axis.text.y = element_text(size=aty),
        axis.title.x = element_text(size=alx),
        axis.title.y = element_text(size=aly))



# Edit Parameters - if plots are the same for both docs, just use one for both ggs

H1 <- "Average Total Interactions of Paid vs Unpaid Posts" # Page/Slide Title goes here
ggWide <- TotalInteractionChart_wide # wider ggplot object goes here
ggTall <- TotalInteractionChart_tall # taller ggplot object goes here
TableTitle <- "Average Total Interactions Table" # Data Table title goes here
DT <- TotalInteraction # Data table object goes here


# Long form text parameters

dataDescription <- 
  "This section shows the mean (labeled on the chart) and distribution Total Interactions on paid versus unpaid posts. In the plot above, we can see the average number of Total.Interactions for Paid posts (via Facebook Ads) is slightly higher than that of Unpaid posts, though the majority of both types of posts have about the same number of Total.Interactions, and both groups have some high outliers skewing their distributions to the right."


dataStrengths <- 
  "Data strengths go here, delete sections as needed."


dataLimitations <- 
  "Data limitations go here."


dataAssumptions <- 
  "Data assumptions go here."




# Add a page to the Word report doc: Title, chart, table, desc., limits, assmps

# DO NOT EDIT THIS, the parameters above will be used

# Notice we are not modifying Report or Presentation, we are making new ones

ReportNew <- Report %>%
  body_add_break() %>% # Adds a page break so this starts on a new page
  body_add_par(value = H1, style = "heading 1") %>% # Title
  body_add_gg(value = ggTall, style = "centered") %>% # ggplot
  body_add_par(value = "") %>% # Adds a space between plot and table
  body_add_par(value = TableTitle, style = "heading 2") %>% # Table title
  body_add_table(value = DT, style = "table_template") %>% # Data table
  body_add_par(value = "") %>% # Adds a space between table and text
  body_add_par(value = "Description", style = "heading 2") %>% 
  body_add_par(value = dataDescription, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Strengths", style = "heading 2") %>% 
  body_add_par(value = dataStrengths, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Limitations", style = "heading 2") %>% 
  body_add_par(value = dataLimitations, style = "Normal") %>%
  body_add_par(value = "") %>% # Adds a space between text and text
  body_add_par(value = "Assumptions", style = "heading 2") %>% 
  body_add_par(value = dataAssumptions, style = "Normal")
  
print(ReportNew, target = "Demo_Report.docx")


# Add another slide to the pptx file: Title and graph/chart
# DO NOT EDIT THIS, the parameters above will be used
PresentationNew <- Presentation %>%
  add_slide(layout = "Title and Content", master = "Facet") %>% 
  ph_with(value = H1, location = ph_location_type(type = "title")) %>% 
  ph_with(ggWide, location = ph_location_type(type = "body"))


print(PresentationNew, target = "Demo_Deck.pptx")


```